---
title: "assignment06"
author: "Jason Grahn"
date: "2/12/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(dplyr)
```

Chapter 11 Page 402: 

#11.9 A

Use the sons data of Table 3.8...
```{r import sons}
sons <- read.table(here::here("assignment06/T3_8_SONS.DAT"))  %>% 
  rename(y1 = V1, # head length
         y2 = V2, # head breadth
         x1 = V3, # head length
         x2 = V4) # head breadth

head(sons, 5)
```

```{r}
#normalize the data
sons.std <-sweep(sons, 2, sqrt(apply(sons,2,var)), FUN="/")
head(sons.std, 5)

#subgroup the data into first and second sons
first.son <- sons.std %>% select(y1, y2)
second.son <- sons.std %>% select(x1, x2)
```

## (a) Find the canonical correlations between ($y_1$,$y_2$) and ($x_1$, $x_1$)·

### the sample correlation matrix
```{r}
S <- cov(sons.std)

#find the sample correlation matrix for either son
R11 <-cor(first.son)
R11
R22 <-cor(second.son)
R22
```

### b) the characteristic equation
```{r}
# correlate column by column for the 4 columns 
        ## first from each
R12 <-c(cor(first.son[,1], 
            second.son[,1]), 
        ## first and second
        cor(first.son[,1], 
            second.son[,2]),
        ## second and first
        cor(first.son[,2], 
            second.son[,1]),
        ## second and second
        cor(first.son[,2], 
            second.son[,2]))

# coerce r12 into a 2x2 matrix
R12 <-matrix(R12, ncol=ncol(R22), byrow=T) 

# R21 is the transpose of R12
R21 <-t(R12)
```

```{r}
# we have to find the eigenvectors and values in order to write the formulas, so this 
# WILL INCLUDE part C

# Finding the E1 and E2 matrices:
E1 <-solve(R11) %*% R12 %*% solve(R22) %*% R21
# and their eigenvector
e1vectors <- round(eigen(E1)$vectors,2)
e1vectors

E2 <-solve(R22) %*% R21 %*% solve(R11) %*% R12
e2vectors <- round(eigen(E2)$vectors,2)
e2vectors
```

So the final formulas are:
  
    u1 =  0.73[y1] + 0.69[y2]
    v1 = -0.68[x1] - 0.71[y2]
    u2 = -0.70[y1] + 0.71[y2]
    v2 = -0.73[x1] + 0.71[y2]

### c) eigenvalues
```{r eigenvalues}
eigen(E1)$values
eigen(E2)$values

#they are the same!
```

### Finally, the cannonical correlation
```{r cannonical correlation}
canon.corr <-sqrt(eigen(E1)$values)
canon.corr
```


#2. Chapter 11 Page 402: #11.9 part (c)

## Test the significance of each canonical correlation.
```{r build degrees of freedom}
# known facts
n <- nrow(sons)
p <- 2
q <- 2
k <- 2

#make $V_H$
vh <- q
vh
#make $V_E$
ve <- n - 1 - q
ve
```

We reject $H_0$ __IF__ $\Lambda_1$ <= $\Lambda_a$.

```{r significance test}
# lambda is 1 - the canonical correlation eigenvalue
test1 <- 1 - canon.corr[1]
test2 <- 1 - canon.corr[2]

# look at Table A.9 for the critical value
# table A.9 shows the critical value for vh = 2 and ve = 22 is .643
CriticalValue <- .643
result1 <- test1 <= CriticalValue
result2 <- test2 <= CriticalValue
```

It is `r result1` that the lambda value (`r test1`) of canonical correlation **1** (`r canon.corr[1]`) is less than or equal to the critical value `r CriticalValue`; so we reject $H_0$ and conclude that the first canonical correlation is significant and could use this relationship for building out multivatiate model.

But it is `r result2` that the lambda value (`r test2`) of canonical correlation **2** (`r canon.corr[2]`) is less than or equal to the critical value `r CriticalValue`; so we fail to reject $H_0$ and conclude that the second canonical correlation is NOT significant.

