---
title: "DA410 Assignment 3"
author: "Jason Grahn"
date: "1/20/2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for piping, mutates, and selects
```

#Homework 3 

##6.27 
Baten, Tack, and Baeder (1958) compared judges' scores on fish prepared by three methods. Twelve fish were cooked by each method, and several judges tasted fish samples and rated each on four variables: $y_{1}$ = aroma, $y_{2}$ = flavor, $y_{3}$ = texture, and $y_{4}$ = moisture. The data are in Table 6.17. Each entry is an average score for the judges on that fish. 

```{r build matrix for 6.27}
fish <- 
    readr::read_table2(file = here::here("/assignment03/T6_17_FISH.DAT"),
                   col_names = FALSE) %>% 
  select(X1:X5) %>% 
  mutate(X1 = factor(X1))

head(fish,10)
```

###(a) 
Compare the three methods using all four MANOVA tests.
```{r build the manova}
manova.data <- manova(cbind(X2 ,X3, X4, X5) ~ X1,
                     data=fish)

broom::tidy(manova.data)
```

#### Wilks' test
```{r wilks fish}
fish.lambda <- summary(manova.data, test = "Wilks")
fish.lambda
```

Wilks' $\Lambda$ is low, so we can reject $H_0$.

#### Pillai test
```{r pillai fish}
fish.pillai <- summary(manova.data, test = "Pillai")
fish.pillai
```

Pillai test is low, so we can reject $H_0$.

#### Roy test
```{r roy fish}
fish.roy <- summary(manova.data, test = "Roy")
fish.roy 

#roy value is 137.168, which we convert...
2.9515/(1+2.9515)
```

We can reject $H_0$ for Roy as well.

#### Hotelling-Lawley test:
```{r hotel fish}
fish.hotel <- summary(manova.data, test = "Hotelling-Lawley")
fish.hotel
```

$H_0$ for Hotelling-Lawley test is also low, so that can be rejected.

```{r fish aov summary}
summary.aov(manova.data)
```

##6.28 
Table 6.18 from Keuls et al. (1984) gives data from a two-way (fixed-effects) MANOVA on snap beans showing the results of four variables: $y_{1}$ = yield earliness, $y_{2}$ = specific leaf area (SLA) earliness, $y_{3}$ = total yield, and $y_{4}$ = average SLA. The factors are sowing date (S) and variety (V).

```{r build matrix for 6.28}
snapbeans <- 
  readr::read_table2(file = here::here("/assignment03/T6_18_SNAPBEAN.DAT"),
                   col_names = FALSE) %>% 
  select(X1:X7) %>% 
  rename(S = X1,
         V = X2,
         pl = X3,
         y1 = X4,
         y2 = X5,
         y3 = X6,
         y4 = X7) %>% 
  mutate(S = factor(S),
         V = factor(V),
         pl = factor(pl))

head(snapbeans, 5)
```


```{r build beans manova }
manova.beans <- manova(cbind(y1 ,y2, y3, y4) ~ S * V,
                       data=snapbeans)

manova.beans
```


###(a)
Test for main effects and interaction using all four MANOVA statistics.

####Wilks’ Test
(this is a kind of test to measure if means are equal)
```{r beans wilks}
beans.lambda <- summary(manova.beans, test = "Wilks")
beans.lambda
```

Wilks $\Lambda$ for S is low, reject $H_0$.
Wilks $\Lambda$ for V is low, reject $H_0$.
Wilks $\Lambda$ for the S*V interaction is low, reject $H_0$.


#### Pillai test
```{r beans pillai}
beans.pillai <- summary(manova.beans, test = "Pillai")
beans.pillai
```

$H_0$ is rejected in Pillai's test as well. 

#### Roy test
```{r beans roy}
beans.roy <- summary(manova.beans, test = "Roy")

#roy values are 137.168, 11.445, and 2.649
S <- 137.168/(1+137.168)
V <- 11.445/(1+11.445)
SV <- 2.649/(1+2.649)

S
V
SV
```

$H_0$ is rejected with Roy's test.

#### Hotelling-Lawley test:
```{r beans hotel}
beans.hotel <- summary(manova.beans, test = "Hotelling-Lawley")
beans.hotel
```

And finally, we reject $H_0$ with Hotelling-Lawley test. 

#Let's try this with tidy? 
https://rdrr.io/github/tidyverse/broom/man/tidy.manova.html is a resource

```{r}
#need to figure out something to put in stats::manova()
bean.manova <- stats::manova(cbind(y1, y2, y3, y4) ~ S*V, snapbeans)

#then can wrap it in tidy() from the "broom" package
roys.tidy.beans <- broom::tidy(bean.manova, 
                              test = "Roy",
                              intercept = FALSE)

roys.tidy.beans

#but why, Jason? 
#because I don't like doing math! So I have R do it for me. 

roys.tidy.beans %>%  #use that tidy tibble
  filter(term != "Residuals") %>%  #filter out residuals because I dont need them
  select(term, roy) %>% #keep the columns I want
  mutate(roys_theta = round((roy / (1 + roy)), 3)) #make roy's theta

#Done!
```



