---
title: "da410_project3"
author: "Jason Grahn"
date: "1/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(MASS)
```

#Problem 1.

## 1. Use admission.csv as a training dataset.

## 2. Train model using LDA by setting admit/not-admit/border with the same probabilities.

## 3. Calculate the misclassfication rate

## 4. Predict students with GPA and SAT score as below.
3.14 	470
3.08 	591
2.08 	641
3.22 	463
```{r build a dataframe of the students to predict}
p_gpa <- c(3.14, 3.08, 2.08, 3.22)
p_gmat <- c(470, 591, 641, 463)
p_student <- c("student1", "student2", "student3", "student4")

predict_me <- table(p_student, p_gpa, p_gmat)
predict_me
```


# Problem 2.

## 1. Use admission.csv as a training dataset.

## 2. Train model using LDA by setting probability of admit is 50% while probability of not admit is 25% and probability of border is 25%.

## 3. Calculate the misclassfication rate

## 4. Predict students with GPA and SAT score as below.
3.14 	470
3.08 	591
2.08 	641
3.22 	463

## Compare differences of the result from problem1.

# Problem 3.

## Explain what is Quadratic Discriminant Analysis (QDA), and use QDA to train the model, discuss if this project can be done better by QDA, why or why not. 