---
title: "final"
author: "jason grahn"
date: "3/16/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


# Problem 1

Show your reason completely and clearly, including all necessary commands and outputs, as
well as interpretations/conclusions.

```{r}
S <- matrix(c(5, 0, 0, 0, 9, 0, 0, 0, 8), byrow = TRUE, nrow = 3, ncol = 3)
S
```

## (a) Find the eigenvalues and eigenvectors of S. 
```{r using covariance S}
# 2. Calculate eigenvectors/eigenvalues
res.eig <- (eigen(S))
res.eig
```

## (b) Show the percent of variance explained.
```{r variance explanation}
R.prop <- round(res.eig$values/(sum(res.eig$values)), 4)
R.prop

round(sum(R.prop), 2)
```

## (c) Decide how many components to retain.

I would keep all three components for this analysis. [[[[explain]]]]


# Problem 2:

The correlation matrix given below arises from the scores of 220 boys in six school subjects: (1) French, (2) English, (3) history, (4) arithmetic, (5) algebra, and (6) geometry. 

## Obtain principal component loadings for three factors.
```{r}
p2 = matrix(c(1, 0.44, 0.41, 0.29, 0.33, 0.25, 0.44, 1, 0.35, 0.35, 0.32, 0.33,
              0.41, 0.35, 1, 0.16, 0.19, 0.18, 0.29, 0.35, 0.16, 1, 0.59, 0.47,
              0.33, 0.32, 0.19, 0.59, 1, 0.46, 0.25, 0.33, 0.18, 0.47, 0.46, 1),
            nrow=6, ncol=6, byrow = TRUE) 
colnames(p2) <- c("French", "English", "History", "Arithmetic", "Algebra", "Geometry")
rownames(p2) <- c("French", "English", "History", "Arithmetic", "Algebra", "Geometry")
#p2

library(psych)
fit <- principal(p2, 
                 nfactors = 3, 
                 rotate = 'none', 
                 covar = FALSE)
fit
```




# Problem 3:
For the Foodstuff Contents data set below, column names are “Food Energy Protein Fat Calcium Iron”:

## (a) Discuss your choice of the number of factors.
## (b) Obtain principal component loadings.
## (c) Calculate percent of variance explained for each factor, plot the factor scores using appropriate plot(s), and decide how many components to retain. 

Show your work completely and clearly, including all necessary commands and outputs, as well as interpretations/conclusions.

```{r}
library(readr)
Problem3_data <- read_csv(here::here("final/Problem3-data.csv"))
head(Problem3_data,3)


```


# Problem 4:
The data below measures in five variables in comparison of normal patients and diabetics:
x1 : glucose intolerance
x2 : insulin response to oral glucose
x3 : insulin resistance
y1 : relative weight
y2 : fasting plasma glucose

## (a) Find the canonical correlation between (y1, y2) and (x1, x2, x3).
## (b) Test the significance of each canonical correlation.
Show your work completely and clearly, including all necessary commands and outputs, as well
as interpretations/conclusions.

```{r}

```


# Problem 5:

The data consists of mental ability test scores of seventh- and eighth-grade children from two
different schools (Pasteur and Grant-White). In our version of the dataset, only 9 out of the
original 26 tests are included. A CFA model that is often proposed for these 9 variables consists of
three latent variables (or factors), each with three indicators:

* a visual factor measured by 2 variables: x1 and x2
* a textual factor measured by 4 variables: x3, x4, x5 and x6
* a speed factor measured by 3 variables: x7, x8 and x9
* a visual factor and a textual factor have zero correlation

## (a) Please draw a figure contains a graphical representation of the three-factor model.
## (b) Please write out the corresponding syntax for specifying this model.

```{r}

```

# Problem 6:
Make a conclusion for this class DA 410, make sure you include the following aspects:
## (a) How many models you have learnt, use 3 to 5 sentences to explain each of them.
## (b) Which one really impressed you when you learnt and why
## (c) Which one is your favorite one and why
## (d) Select two models out, make a comparison. Show the differences and similarities between
them.
## (e) If you will build up a project to solve some real problem using one of them, which one you
would like to you, and what kind of project you will like to build. (200 words)

