---
title: "final"
author: "jason grahn"
date: "3/16/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(psych)
library(readr)
library(magrittr)
library(dplyr)
library(yacca)
library(lavaan) 
library(semPlot)
```

# Problem 1

Show your reason completely and clearly, including all necessary commands and outputs, as
well as interpretations/conclusions.

```{r}
S <- matrix(c(5, 0, 0, 0, 9, 0, 0, 0, 8), byrow = TRUE, nrow = 3, ncol = 3)
S
```

## (a) Find the eigenvalues and eigenvectors of S. 
```{r using covariance S}
# 2. Calculate eigenvectors/eigenvalues
res.eig <- (eigen(S))
res.eig
```

## (b) Show the percent of variance explained.
```{r variance explanation}
R.prop <- round(res.eig$values/(sum(res.eig$values)), 4)
R.prop

round(sum(R.prop), 2)

0.4091 + 0.3636
```

## (c) Decide how many components to retain. 

Utilizing only the first factor accounts for 40% of variance, which isn't enough. Using the first two provides 77% of variance explained, which would be acceptible for any business setting, but doesn't seem sufficient for our causes. I would use all three factors  I would keep all three components for this analysis as the third factor contains nearly 23% of explained variance, which is significant enough to include; and all three add up to 100% of explained variance. 

# Problem 2:

The correlation matrix given below arises from the scores of 220 boys in six school subjects: (1) French, (2) English, (3) history, (4) arithmetic, (5) algebra, and (6) geometry. 

## Obtain principal component loadings for three factors.
```{r}
p2 = matrix(c(1, 0.44, 0.41, 0.29, 0.33, 0.25, 0.44, 1, 0.35, 0.35, 0.32, 0.33,
              0.41, 0.35, 1, 0.16, 0.19, 0.18, 0.29, 0.35, 0.16, 1, 0.59, 0.47,
              0.33, 0.32, 0.19, 0.59, 1, 0.46, 0.25, 0.33, 0.18, 0.47, 0.46, 1),
            nrow=6, ncol=6, byrow = TRUE) 
colnames(p2) <- c("French", "English", "History", "Arithmetic", "Algebra", "Geometry")
rownames(p2) <- c("French", "English", "History", "Arithmetic", "Algebra", "Geometry")
#p2

fit <- principal(p2, 
                 nfactors = 3, 
                 rotate = 'none', 
                 covar = FALSE)
fit
```

`PC1` shows us that all variables are positively strong. `PC2` shows us that History is strongly positive, while the others start to degrade in weakness. `PC3` shows us that History maintains that strong positive relationship, and English has a strong negative relationship, while the rest of the variables seem insignificant. 

# Problem 3:

For the Foodstuff Contents data set below, column names are “Food Energy Protein Fat Calcium Iron”:

## (a) Discuss your choice of the number of factors.
## (b) Obtain principal component loadings.
## (c) Calculate percent of variance explained for each factor, plot the factor scores using appropriate plot(s), and decide how many components to retain. 

Show your work completely and clearly, including all necessary commands and outputs, as well as interpretations/conclusions.

```{r}
Problem3_data <- read_csv(here::here("final/Problem3-data.csv"))
head(Problem3_data,3)

fit <- principal(Problem3_data[2:6], 
                 nfactors = 4, 
                 rotate = 'none', 
                 covar = FALSE)
fit

# scale, center, and apply PCA 
probe.pca <- prcomp(Problem3_data[2:6],
                    center = TRUE,
                    scale. = TRUE) 
                    
screeplot(probe.pca, type = "l", main = "Screeplot of the PCs")
abline(h = 0.5, 
       col="red", 
       lty=5)
legend("topright", 
       legend=c("Eigenvalue = .5"),
       col=c("red"), 
       lty=5, 
       cex=0.6)
```


# Problem 4:
The data below measures in five variables in comparison of normal patients and diabetics:

* x1 : glucose intolerance
* x2 : insulin response to oral glucose
* x3 : insulin resistance
* y1 : relative weight
* y2 : fasting plasma glucose

## (a) Find the canonical correlation between (y1, y2) and (x1, x2, x3).

```{r}
Problem4_data <- read_csv(here::here("final/problem4-data.csv"))
head(Problem4_data,3)

normals <- Problem4_data %>% select(y1, y2)
diabetics <- Problem4_data %>% select(x1, x2, x3)

cca_output <- cca(normals, diabetics)
cca_output$corr
```

## (b) Test the significance of each canonical correlation.

```{r}
F.test.cca(cca_output)
```

# Problem 5:

The data consists of mental ability test scores of seventh- and eighth-grade children from two
different schools (Pasteur and Grant-White). In our version of the dataset, only 9 out of the
original 26 tests are included. A CFA model that is often proposed for these 9 variables consists of
three latent variables (or factors), each with three indicators:

* a visual factor measured by 2 variables: x1 and x2
* a textual factor measured by 4 variables: x3, x4, x5 and x6
* a speed factor measured by 3 variables: x7, x8 and x9
* a visual factor and a textual factor have zero correlation

## (a) Please draw a figure contains a graphical representation of the three-factor model.
```{r}
pgwmodel <- ' visual  =~ x1 + x2
              textual =~ x3 + x4 + x5 + x6 
              speed   =~ x7 + x8 + x9 
              visual  ~~ 0*textual ' 

fit <- sem(pgwmodel, 
           data = HolzingerSwineford1939)

semPaths(fit, 
         "std",
         title = "FALSE",
         edge.color = "blue",
         color = "light blue",
         rotation = 2,
         layout = "tree2")
```

## (b) Please write out the corresponding syntax for specifying this model.

*$visual$ =~ $x1$ + $x2$
*$textual$ =~ $x3$ + $x4$ + $x5$ + $x6$  
*$speed$ =~ $x7$ + $x8$ + $x9$
*$visual$ ~~ $0textual$


# Problem 6:
Make a conclusion for this class DA 410, make sure you include the following aspects:
## (a) How many models you have learnt, use 3 to 5 sentences to explain each of them.
## (b) Which one really impressed you when you learnt and why
## (c) Which one is your favorite one and why
## (d) Select two models out, make a comparison. Show the differences and similarities between
them.
## (e) If you will build up a project to solve some real problem using one of them, which one you
would like to you, and what kind of project you will like to build. (200 words)

