---
title: "da410_project2"
author: "Jason Grahn"
date: "1/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

# Part 1: 

## The problem statement

Use Hotelling’s T^2 test to test for a difference in the mean score vector of the boys and the mean vector of the girls. Make sure you include clear command lines and relevant output/results with hypotheses, test result(s) and conclusion(s)/interpretation(s).

## The data

Download testscoredata.txt and read it in R

```{r import data}
scores <- 
  readr::read_table2(file = here::here("/project2/testscoredata.txt"),
                   col_names = TRUE) %>% 
  mutate(subject = factor(subject),
         sex = factor(sex))

head(scores,5)
```

## The hypothesis

> $H_0$: The mean scores between sexes is equal ( $\mu_o = \mu_1$).

> $H_a$: The mean scores between sexes is not equal ( $\mu_o \neq \mu_1$). 

```{r setting up the manova binding for scores}
#build the manova binding
scores.manova <- stats::manova(cbind(math, reading) ~ sex, scores)
scores.manova
```

We see that *resp 1* has a mean value of 19.976, while *resp 2* has a mean value of 3.768. Our initial impression is that the mean values are not equal. 

## The test 
```{r testing scores with hl}
#use the HL method to test results.
tidy.score.hotel <- broom::tidy(scores.manova, 
                              test = "Hotelling-Lawley",
                              intercept = FALSE)
```

##The result: 
```{r show results for scores hl test}
tidy.score.hotel
```

## The conclusion

With a $p-value$ of `r tidy.score.hotel$p.value[1]`, we can reject $H_0$ and conclude that the mean scores between the sexes is not equal.

________
________
________

#Part 2: 

## The problem statement

Suppose we have gathered the following data on female athletes in three sports. The measurements we have made are the athletes' heights and vertical jumps, both in inches. The data are listed as (height, jump) as follows:

    Basketball Players: (66, 27), (65, 29), (68, 26), (64, 29), (67, 29)
    Track Athletes:     (63, 23), (61, 26), (62, 23), (60, 26)
    Softball Players:   (62, 23), (65, 21), (63, 21), (62, 23), (63.5, 22), (66, 21.5)

 * Use R to conduct the MANOVA F-test using Wilks' Lambda to test for a difference in (height, jump) mean vectors across the three sports. Make sure you include clear command lines and relevant output/results with hypotheses, test result(s) and conclusion(s)/interpretation(s)
 * State the assumptions of your test and check to see whether assumptions are met. Do you believe your inference is valid? Why or why not?
 * Use R to examine the sample mean vectors for each group. Make sure you include clear command lines and relevant output/results. Also comment on the differences among the groups in terms of the specific variables.

## The data

Download testscoredata.txt and read it in R

```{r import matrix data}
#imput data
sport  <- as.factor(c('B','B','B','B','B','T','T','T','T','S','S','S','S','S','S'))
height <- c(66,65,68,64,67,63,61,62,60,62,65,63,62,63.5,66)
jump   <- c(27,29,26,29,29,23,26,23,26,23,21,21,23,22,21.5)

#make a table
sports <- tibble(sport, height, jump)
head(sports, 5)
```

## The hypothesis

> $H_0$: The mean difference between sports is equal ( $\mu_o = \mu_1$).

> $H_a$: The mean difference between sports is not equal ( $\mu_o \neq \mu_1$). 

```{r setting up the manova binding}
sports.manova <- stats::manova(cbind(height, jump) ~ sport, sports)
sports.manova
```

## The test 
```{r testing  with Wilks lambda}

```

##The result: 
```{r show results for  Wilks lambda}

```

Wilks' $\Lambda$ score is 

## The conclusion

With a $p-value$ of `r tidy.score.hotel$p.value[1]`, we can reject $H_0$ and conclude that the mean scores between the sexes is not equal.

