---
title: "project5"
author: "Jason Grahn"
date: "2/16/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r package load, message=FALSE, warning=FALSE}
library(CCA)
library(tidyverse)
library(GGally)
library(yacca)
```

#Use R to solve Chapter 11 Page 402: #11.9
Make sure you include the commands and outputs, as well as the interpretations of the outputs.

11.9 Use the sons data of Table 3.8.
```{r import sons data, message=FALSE}
sons <- read.table(here::here("assignment06/T3_8_SONS.DAT"))  %>% 
  rename(y1 = V1, # head length
         y2 = V2, # head breadth
         x1 = V3, # head length
         x2 = V4) # head breadth

head(sons, 5)

ggpairs(sons)
```

```{r subgrouping sons}
#subgroup the data into first and second sons
first.son <- sons %>% select(y1, y2)
second.son <- sons %>% select(x1, x2)
```

```{r build the analysis output}
cca_output <- cca(first.son, second.son)
cc_output <- cc(first.son, second.son)
```

## (a) Find the canonical correlations between ($y_1$,$y_2$) and ($x_1$, $x_1$)·

```{r}
cca_output$corr
```

$r_1$ is `r cca_output$corr[1]` and $r_2$ is `r cca_output$corr[2]`. This shows us that the canonical correlation between [one thing] and [the other thing] is positive and strong, while the canonical correlation between [something else] and [one more thing], while positive, is quite weak.

#(b) Find the standardized coefficients for the canonical variates.
```{r}
cca_output[3:4]
```

#(c) Test the significance of each canonical correlation.
```{r}
F.test.cca(cca_output)
```

The output of the F-test shows us the 