---
title: "da410_assignment5"
author: "Jason Grahn"
date: "2/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

Chapter 9  Page 336-337:#9.7 (a)(b)(c)
9.7 c, don't need to generate classification table, just use nearest neighbor method to assign 3rd vector of both two matrix into group1 or group2. For the detail, look at the example.

You can use the data1.txt and data2.txt to read into data.

Calculate step by step (you can use R to do matrix calculation, DO NOT use knn functions), make sure to provide intermediate steps. 

#9.7 Do a classification analysis on the beetle data in Table 5.5 as follows:
```{r import beetle data}
beetles <- read.table(here::here("assignment05/T5_5_FBEETLES.DAT")) %>% 
  rename(experiment_number = V1,
         species = V2,
         y1 = V3,
         y2 = V4,
         y3 = V5,
         y4 = V6) %>% 
  mutate(species = factor(species, 
                          levels = c(1, 2)))

head(beetles,5)
```


##(a) Find the cutoff point THEN the classification function. 

Cutoff point = $\frac{1}{2}(\bar{z_1}+\bar{z_2})$
Classification function = $z = (\bar{y}_1 - \bar{y}_2)' * S^{-1}_{p1}$
```{r}
# group by species
beet1 <- subset(x = beetles, subset = (species == 1), y1:y4)
beet2 <- subset(x = beetles, subset = (species == 2), y1:y4)

# find group means for all variables for each species
mean1 <- apply(X = beet1, FUN = mean, MARGIN = 2)
mean1
mean2 <- apply(X = beet2, FUN = mean, MARGIN = 2)
mean2

# find covariance matrix for each species
var1 <- var(beet1)
var1
var2 <- var(beet2)
var2

# take a pause here. Check for equal covariances between groups (species)
     # If ==, then proceed; if != then we need a different procedure
# heplots::boxM(cbind(y1, y2, y3, y4) ~ species, data = beetles)
# can't use heplots because samplesize is too small. using hotelling 
beetles.manova <- stats::manova(cbind(y1, y2, y3, y4) ~ species, 
                                data = beetles)

tidy.beetles.manova <- broom::tidy(beetles.manova, 
                              test = "Hotelling-Lawley",
                              intercept = FALSE)
tidy.beetles.manova 
#### because hotelling t-test has p-value <0.001, we conclude equal covariance
```

```{r}
# find _pooled_ sample variance for each species
#length of each set
length1 <- nrow(beet1)
length2 <- nrow(beet2)

#variance of each set
pvar1 <- (length1 - 1) * var1
pvar2 <- (length2 - 1) * var2 

#use length and variance to get pooled variance
sp1 <- 1 / (length1 + length2 - 2) * (pvar1 + pvar2)
sp1
```

```{r}
# find the cutoff point (this is half the difference between group means) 
cutoff <- .5*(mean1 - mean2) %*% solve(sp1) %*% (mean1 + mean2)
cutoff <- round(cutoff,2)
cutoff
```

```{r}
# find the classification function
species_prediction <- apply(beetles[,3:6], 1, function(y) {
 z <- (mean1 - mean2) %*% solve(sp1) %*% y })
``` 

##(b) Find the classification table using the linear classification function in part (a).
```{r}
beetles <- beetles %>% 
  #bring the species prediction data into the data frame
  mutate(species_prediction = species_prediction,
         #bring the cutoff data into the data frame
         cutoff = cutoff,
         #use the cutoff to determine the classification
         classification = if_else(species.prediction >= cutoff, 1, 2))

table(beetles$species, beetles$classification)
```

##(c) Describe how you use KNN thought to solve it instead of writing out R code or doing any real calculation.

Start by selecting a data point in the first group. 
Calculating a "normalized" distance from data point 1 in Group1 to all other points in both groups. This is _typically_ Euclidian distance. 
Select the _k_-value. If the number of groups is even, this has to be an odd number greater than the amount of groups. If the number of groups is odd, you can use this number.
The algorithm will determine who are the __nearest neightbors__ to a new data point. Whichever group has the highest number of neighbors to the new data point is considered the majority group. This is the group the new data point is assigned to. 