---
title: "da410_assignment5"
author: "Jason Grahn"
date: "2/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
```

Chapter 9  Page 336-337:#9.7 (a)(b)(c)
9.7 c, don't need to generate classification table, just use nearest neighbor method to assign 3rd vector of both two matrix into group1 or group2. For the detail, look at the example.

You can use the data1.txt and data2.txt to read into data.

Calculate step by step (you can use R to do matrix calculation, DO NOT use knn functions), make sure to provide intermediate steps. 

#9.7 Do a classification analysis on the beetle data in Table 5.5 as follows:
```{r import beetle data}
beetles <- read.table(here::here("assignment05/T5_5_FBEETLES.DAT")) %>% 
  rename(experiment_number = V1,
         species = V2,
         y1 = V3,
         y2 = V4,
         y3 = V5,
         y4 = V6) %>% 
  mutate(species = factor(species, 
                          levels = c(1, 2)))

head(beetles,5)
```


##(a) Find the cutoff point THEN the classification function. 

Cutoff point = $\frac{1}{2}(\bar{z_1}+\bar{z_2})$
Classification function = $z = (\bar{y}_1 - \bar{y}_2)' * S^{-1}_{pl}$
```{r}
# group by species
beet1 <- subset(x = beetles, subset = (species == 1), y1:y4)
beet2 <- subset(x = beetles, subset = (species == 2), y1:y4)

# find group means for all variables for each species
mean1 <- apply(X = beet1, FUN = mean, MARGIN = 2)
mean1
mean2 <- apply(X = beet2, FUN = mean, MARGIN = 2)
mean2

# find covariance matrix for each species
var1 <- var(beet1)
var1
var2 <- var(beet2)
var2

# take a pause here. Check for equal covariances between groups (species)
     # If ==, then proceed; if != then we need a different procedure
# heplots::boxM(cbind(y1, y2, y3, y4) ~ species, data = beetles)
# can't use heplots because samplesize is too small. using hotelling 
beetles.manova <- stats::manova(cbind(y1, y2, y3, y4) ~ species, 
                                data = beetles)

tidy.beetles.manova <- broom::tidy(beetles.manova, 
                              test = "Hotelling-Lawley",
                              intercept = FALSE)
tidy.beetles.manova 
#### because hotelling t-test has p-value <0.001, we conclude equal covariance
```

```{r}
# find _pooled_ sample variance for each species
#length of each set
length1 <- nrow(beet1)
length2 <- nrow(beet2)

#variance of each set
pvar1 <- (length1 - 1) * var1
pvar2 <- (length2 - 1) * var2 

#use length and variance to get pooled variance
sp1 <- 1 / (length1 + length2 - 2) * (pvar1 + pvar2)
sp1

# find the cutoff point (this is half the difference between group means) 
cutoff <- .5*(mean1 - mean2) %*% solve(sp1) %*% (mean1 + mean2)
round(cutoff,2)
```

```{r}
# find the classification function

``` 

##(b) Find the classification table using the linear classification function in part (a).
```{r}

```

##(c) Find the classification table using the nearest neighbor method.
```{r}

```

